\Huge ACM and Climate Change

### 2018-12-11 main.tex
Finished first draft

\newcommand{\bcp}[1]{\ifdraft{\bf #1}\fi}

@@

Why ACM? Because one of ACM’s main activities is organizing international
conferences, and these produce significant emissions of greenhouse gases,
particularly from air travel. For example, a single round-trip flight from
Philadelphia to Paris emits the equivalent of about 2 tons of carbon dioxide
(CO2e) per passenger. Compare this against the fact that the average US
resident emits about 16.5 tons per year, while the average European emits
about 7 tons [10].  Moreover, these emissions have no near-term
technological fix because jet fuel is difficult to replace with renewable
energy sources [7]. So reducing the CO2e footprint of conferences is going
to require a reduction in the footprint of the attendees’ air travel.

We are the members of an {\em ad hoc} Committee on Climate Change within
SIGPLAN (ACM’s Special Interest Group on Programming Languages) that was
formed in 2016 to consider possible strategies for mitigating the emissions
of our conferences [3]. Among the ones we consider most attractive are:
making it easier to use carbon offsets to mitigate aviation emissions for
conferences, choosing conference locations with an eye to minimizing
emissions, co-locating or merging conferences, virtualizing program
committees, and supporting remote participation.

Buying carbon offsets to reduce the impact of air travel is an ``easy win''
[4].  The idea of carbon offsets is to pay somebody to plant trees, capture
methane (from landfills or cattle herds, for example), buy cookstoves to
replace inefficient open-fire cooking in poor communities, or carry out
other actions that remove greenhouse gases from the air or prevent their
emission. At the moment, the average air travel to an international
conference can be offset by paying around \$30 [4].  ACM could streamline
the purchase of such offsets and make offsetting the default by folding them
into its conference registration fees.

Unfortunately, carbon offsetting is only a short-term mitigation, since the
potential supply of offsets is far less than the potential demand [11]. We
must also begin significantly reducing emissions, which means reducing the
distance traveled by the sum of all conference participants.

One approach to doing so is to choose conference locations that are close to
the center of mass of their potential participants.  For example, SIGGRAPH
(the Special Interest Group on Graphics) locates in Los Angeles every other
year---an easy call as this is near an international airport and a major hub
of the worldwide entertainment industry. For other research areas, the
effects of location choice are less obvious. For example,
Figure~\ref{fig:SIGPLAN-confs} plots the 
locations of the four major SIGPLAN conferences over the past five years,
colored according to the estimated average carbon footprint per
participant. The figure shows that there were significant differences across
conferences and across locations---indeed, it is a factor of two between the
lowest and highest. Why is this? One reason might be that the locus of a
conference’s population is closer to one location than another, and changing
locations reduces the total cost. Another might be the presence of a local
population that will only attend if the conference is very close to
them. Our preliminary analysis suggests it may be both. The larger question
is how to use this kind of information to choose the best locations over
time, for a diffuse population.

@@

We can attend some conferences online, saving both emissions and travel
time.  Indeed, SIGPLAN already livestreams all its four main conferences,
while SIGACCESS and SIGCHI have experimented with telepresence robots at
their main conferences [8,9]. Journal-first initiatives, which make
conference attendance optional, could also help.  Looking further ahead,
some communities are already experimenting with fully virtual conferences
[6].  ACM members can contribute to the development of technologies for
connecting people effectively at a distance.

@@

The problem of climate change is too large and too urgent to leave to world
leaders: organizations at every level must play their part in dealing with
their own contributions, raising awareness, and establishing new ways of
doing business in the lower-carbon future that is coming. For ACM, this
especially means developing strategies for reducing the climate impact of
conference-related travel. You can help in a number of ways: 
\begin{itemize}
\item {\bf Get involved} in raising awareness about this issue in the
organizations you are a part of.  For example, ask the organizers of the ACM
conferences you attend what they are doing about reducing or offsetting
emissions from participants’ travel.
\item {\bf Analyze} your organization’s carbon-emitting activities to inform a
discussion of how best to reduce its footprint.
\item {\bf Connect} with others that are engaging with these issues---in
your workplace, in your other communities, and within ACM (e.g., on the
acm-climate mailing list [5]).
\end{itemize}

@@

\section*{More References (to be bibtexified)}
\begin{itemize}
\item 
[3] Engaging with Climate Change: Possible Steps for SIGPLAN. Preliminary Report of the SIGPLAN Climate Committee.  Michael Hicks,​ ​Crista Lopes, and​ ​Benjamin C. Pierce. Version 1.2, June 2018. http://www.cis.upenn.edu/~bcpierce/papers/sigplan-climate-report.pdf
\item [4] Carbon Offsets: An Overview for Scientific Societies.  Richard Kim and ​Benjamin C. Pierce. Version 1.2 June 24, 2018.  http://www.cis.upenn.edu/~bcpierce/papers/carbon-offsets.pdf
\item[5] acm-climate mailing list, https://groups.google.com/forum/\#!forum/acm-climate
[6] OpenSimulator Community Conference, https://conference.opensimulator.org/2018/
\item [7] Why Aren’t There Electric Airplanes Yet? Venkat Viswanathan, Shashank Sripad, and William Leif Fredericks. November 2018. https://www.universal-sci.com/headlines/2018/11/30/-why-arent-there-electric-airplanes-yet
\item [8] CHI 2018, Remote Attendance, https://chi2018.acm.org/attending/remote-attendance/
\item [9] Neustaedter, C., Tang, J., Singhal, S., Pan, R., Heshmat, Y., and Forghani, A. (2017) Telepresence Attendance at the ACM CSCW 2016 Conference, Connections Lab Technical Report 2017-0419-01, Simon Fraser University.
\item [10] CO2 emissions (metric tons per capita). The World Bank Data Bank, using data from the Carbon Dioxide Information Analysis Center, Environmental Sciences Division, Oak Ridge National Laboratory, Tennessee, United States. Accessed December 2018. https://data.worldbank.org/indicator/EN.ATM.CO2E.PC
\item [11]  Supply and sustainability of carbon offsets and alternative fuels for international aviation, by
Rob Bailis, Derik Broekhoff, and Carrie M. Lee.  Stockholm Environment Institute, Working Paper 2016-03. https://www.sei.org/publications/carbon-offsets-aviation/			
\end{itemize}

\ifdraft
\newpage
\section*{Permissions for figures}

[1] Creative Commons license
[2] ``Reproduction of a limited number of figures or short excerpts of IPCC material is authorized free of charge and without formal written permission provided that the original source is properly acknowledged, with mention of the name of the report, the publisher and the numbering of the page(s) or the figure(s).''
[3] Authors

\section*{Things to do}

\paragraph*{Final things to do:}
\begin{itemize}
\item 
Check where we’ve used ``carbon'' (or CO2) and where ``greenhouse gases''.  Gloss CO2 or CO2e as appropriate.
\end{itemize}

\paragraph*{Discussions:}
\begin{itemize}
\item 
About Crista’s figure:
BCP: Would be great to supplement with a more refined analysis of registrations for each of these conferences; then we could add a "center of mass" dot for each of the conferences to the map. This would still not tell the whole story, since it's possible (likely) that the registration totals would have been different if the conferences had been located differently. But it would help tell the story.
\item 
Also about Crista’s figure:
\begin{itemize}
\item 
BCP: We should quantify what the red and blue dots mean -- i.e., give readers an idea what overall difference a change of location would make to carbon footprint.
\item Mike: The dot colors mean per-capita footprint. Could we make their size represent the total footprint? Far-flung places might be high per-capita, but lower overall.
\item Crista: The order by total footprint is very similar to the normalized one. Here are the 5 top offenders by that criteria: ICFP-Japan, POPL-LA, POPL-SD, SPLASH-Vancouver, PLDI-Barcelona. There are some changes in the order of the conferences in the middle. I hesitated which plot to draw (each has positives and negatives) but decided this one would be less problematic. I don't think we have space for two, and overlaying the 2 data series will be confusing...
\item BTW, there's a LOT more that can be said about this data... But I think that will distract from the main message of this article. We need to chose whether this 2-page article is about the data analysis of our conferences or about a call for action. Maybe we can write a data analysis paper, upload it on arXiv, and link to it from this article?
\item
For example: POPL-Paris was the absolute winner by all criteria: it had the highest number of participants, the second-lowest normalized carbon footprint, and was mid-point in total carbon footprint, even with all those participants.
This is just one tidbit; there are many more interesting facts in this data.
\end{itemize}
\item 
From Michael Coblenz:
What about research on energy-efficient computing? The impact of this research might be far more significant than that of mitigating our direct carbon emissions, and this is more in line with the core skills of the community.
Also, what about research on telepresence and virtual meetings? There’s obviously a lot of work in this area, but we may need more in order to (a) let us effectively trade off options, e.g. virtual PC meeting vs in-person; (b) improve the state of the art to make remote meetings more effective.
Consider also discussing the diversity implications on travel. Reducing travel expectations might make the field more attractive to women, who may face higher social expectations to stay home (and are more likely to be single parents). Surely there is a good article to cite here. I found https://journals.sagepub.com/doi/abs/10.1177/0950017006066999 but maybe there is something better. I’m looking…
\item 
Consider citing the new IEA report as another piece of evidence that action is needed
\item 
From Sophia: Does the article need an abstract to "set the reader's expectations"? I expected that there would be a list of actions in the end, or recommended resolutions. Instead, the article is more about what some communities do, and where individuals may be involved.  Perhaps say: "In this article we give a short overview of how different SCM communities tackle this problem, what thought are relevant to it, and conclude with suggestions on how to get involved". It breaks the flow, but it clarifies what the article is about.
\item 
Is citation 10 CO2 or CO2e?
\item 
Is there a ``more authoritative'' source for price of carbon offsets than [4]?
\end{itemize}
\fi

>>>>

\newcommand{\bcp}[1]{\ifdraft{\bf #1}\fi}
\newcommand{\COtwoE}{CO$_2$e}

@@

Why ACM? Because one of ACM’s main activities is organizing international
conferences, and these produce significant emissions of greenhouse gases,
particularly from air travel. For example, a single round-trip flight from
Philadelphia to Paris emits the equivalent of about 2 tons of carbon dioxide
(\COtwoE) per passenger. Compare this against the fact that the average US
resident emits about 16.5 tons per year, while the average European emits
about 7 tons~\cite{emissions}.  Moreover, these emissions have no near-term
technological fix because jet fuel is difficult to replace with renewable
energy sources~\cite{elec-air}. So reducing the \COtwoE{} footprint of conferences is going
to require a reduction in the footprint of the attendees’ air travel.

We are the members of an {\em ad hoc} Committee on Climate Change within
SIGPLAN (ACM’s Special Interest Group on Programming Languages) that was
formed in 2016 to consider possible strategies for mitigating the emissions
of our conferences \cite{ClimateCommitteReport}. Among the ones we find most
attractive are: making it easier to use carbon offsets to mitigate aviation
emissions for conferences, choosing conference locations with an eye to
minimizing emissions, co-locating or merging conferences, virtualizing
program committees, and supporting remote participation.

Buying carbon offsets to reduce the impact of air travel is an ``easy win''
\cite{CarbonOFfsetReport}.  The idea of carbon offsets is to pay somebody to plant trees, capture
methane (from landfills or cattle herds, for example), buy cookstoves to
replace inefficient open-fire cooking in poor communities, or carry out
other actions that remove greenhouse gases from the air or prevent their
emission. At the moment, the average air travel to an international
conference can be offset by paying around \$30 \cite{CarbonOFfsetReport}.
ACM could streamline 
the purchase of such offsets and make offsetting the default by folding them
into its conference registration fees.

Unfortunately, carbon offsetting is only a short-term mitigation, since the
potential supply of offsets is far less than the potential demand~\cite{SEI-Report}. We
must also begin significantly reducing emissions, which means reducing the
distance traveled by the sum of all conference participants.

One approach to doing so is to choose conference locations that are close to
the center of mass of their potential participants.  For example, SIGGRAPH
(the Special Interest Group on Graphics) locates in Los Angeles every other
year---an easy call as this is near an international airport and a major hub
of the worldwide entertainment industry. For other research areas, the
effects of location choice are less obvious. For example,
Figure~\ref{fig:confs} plots the 
locations of the four major SIGPLAN conferences over the past five years,
colored according to the estimated average carbon footprint per
participant. The figure shows that there were significant differences across
conferences and across locations---indeed, it is a factor of two between the
lowest and highest. Why is this? One reason might be that the locus of a
conference’s population is closer to one location than another, and changing
locations reduces the total cost. Another might be the presence of a local
population that will only attend if the conference is very close to
them. Our preliminary analysis suggests it may be both. The larger question
is how to use this kind of information to choose the best locations over
time, for a diffuse population.

@@

We can attend some conferences online, saving both emissions and travel
time.  Indeed, SIGPLAN already livestreams all its four main conferences,
while SIGACCESS and SIGCHI have experimented with telepresence robots at
their main conferences~\cite{CHI-remote,Neustaedter}. Journal-first
initiatives, which make conference attendance optional, could also help.
Looking further ahead, some communities are already experimenting with fully
virtual conferences~\cite{OS}.  ACM members can contribute to the
development of technologies for connecting people effectively at a distance.

@@

The problem of climate change is too large and too urgent to leave to world
leaders: organizations at every level must play their part in dealing with
their own contributions, raising awareness, and establishing new ways of
doing business in the lower-carbon future that is coming. For ACM, this
especially means developing strategies for reducing the climate impact of
conference-related travel. You can help in a number of ways: 
\begin{itemize}
\item {\bf Get involved} in raising awareness about this issue in the
organizations you are a part of.  For example, ask the organizers of the ACM
conferences you attend what they are doing about reducing or offsetting
emissions from participants’ travel.
\item {\bf Analyze} your organization’s carbon-emitting activities to inform a
discussion of how best to reduce its footprint.
\item {\bf Connect} with others that are engaging with these issues---in
your workplace, in your other communities, and within ACM (e.g., on the
acm-climate mailing list~\cite{acm-climate}).
\end{itemize}

@@

\iflater
\newpage
\section*{Permissions for figures}

[1] Creative Commons license

[2] ``Reproduction of a limited number of figures or short excerpts of IPCC material is authorized free of charge and without formal written permission provided that the original source is properly acknowledged, with mention of the name of the report, the publisher and the numbering of the page(s) or the figure(s).''

[3] Authors

\section*{More Discussions:}
\begin{itemize}
\item 
About Crista’s figure:
BCP: Would be great to supplement with a more refined analysis of registrations for each of these conferences; then we could add a "center of mass" dot for each of the conferences to the map. This would still not tell the whole story, since it's possible (likely) that the registration totals would have been different if the conferences had been located differently. But it would help tell the story.
\item 
Also about Crista’s figure:
\begin{itemize}
\item 
BCP: We should quantify what the red and blue dots mean -- i.e., give readers an idea what overall difference a change of location would make to carbon footprint.
\item Mike: The dot colors mean per-capita footprint. Could we make their size represent the total footprint? Far-flung places might be high per-capita, but lower overall.
\item Crista: The order by total footprint is very similar to the normalized one. Here are the 5 top offenders by that criteria: ICFP-Japan, POPL-LA, POPL-SD, SPLASH-Vancouver, PLDI-Barcelona. There are some changes in the order of the conferences in the middle. I hesitated which plot to draw (each has positives and negatives) but decided this one would be less problematic. I don't think we have space for two, and overlaying the 2 data series will be confusing...
\item BTW, there's a LOT more that can be said about this data... But I think that will distract from the main message of this article. We need to chose whether this 2-page article is about the data analysis of our conferences or about a call for action. Maybe we can write a data analysis paper, upload it on arXiv, and link to it from this article?
\item
For example: POPL-Paris was the absolute winner by all criteria: it had the highest number of participants, the second-lowest normalized carbon footprint, and was mid-point in total carbon footprint, even with all those participants.
This is just one tidbit; there are many more interesting facts in this data.
\end{itemize}
\item 
From Michael Coblenz:
What about research on energy-efficient computing? The impact of this research might be far more significant than that of mitigating our direct carbon emissions, and this is more in line with the core skills of the community.
Also, what about research on telepresence and virtual meetings? There’s obviously a lot of work in this area, but we may need more in order to (a) let us effectively trade off options, e.g. virtual PC meeting vs in-person; (b) improve the state of the art to make remote meetings more effective.
Consider also discussing the diversity implications on travel. Reducing travel expectations might make the field more attractive to women, who may face higher social expectations to stay home (and are more likely to be single parents). Surely there is a good article to cite here. I found https://journals.sagepub.com/doi/abs/10.1177/0950017006066999 but maybe there is something better. I’m looking…
\item 
Consider citing the new IEA report as another piece of evidence that action is needed
\item 
From Sophia: Does the article need an abstract to "set the reader's expectations"? I expected that there would be a list of actions in the end, or recommended resolutions. Instead, the article is more about what some communities do, and where individuals may be involved.  Perhaps say: "In this article we give a short overview of how different SCM communities tackle this problem, what thought are relevant to it, and conclude with suggestions on how to get involved". It breaks the flow, but it clarifies what the article is about.
\item 
Is the citation for emissions by region CO2 or CO2e?
\item 
Is there a ``more authoritative'' source for price of carbon offsets than
our carbon report?
\end{itemize}
\item \item 
Check where we’ve used ``carbon'' (or CO2) and where ``greenhouse gases''.  Gloss CO2 or CO2e as appropriate.
\fi
